---
title: "Result for Homogenizing Model"
author: "Yixuan Yang"
date: '2022-07-05'
output: pdf_document
header-includes:
  - \usepackage{caption}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
rm(list = ls()) # clean work space
load('../Rdata/Fish.Rdata')
library(tidyverse)
library(kableExtra)
```

**Number of data: 39**

\
\

```{r, echo=FALSE}
p1 = data.frame(IGC1_Full_Homo$tau, IGC1_Full$tau)
ggplot(data = p1, aes(x = IGC1_Full.tau, y = IGC1_Full_Homo.tau)) +
  geom_point() +
  geom_abline() +
  labs(title = "Homogenizing tau v.s. original tau") +
  xlab("original tau") +
  ylab("homogenizing tau")
```

Plot 1: comparing tau value between homogenizing model and Xiang's original model

**Homogenizing $\tau$ < original $\tau$:**

```{r,echo=FALSE}
rownames(IGC1_Full)[which(p1[,1] < p1[,2])]
```

```{r, echo=FALSE}
p2 = data.frame(IGC1_Full_Homo$homogenizing_omega, IGC1_Full_Homo$omega)
ggplot(data = p2, aes(x = IGC1_Full_Homo.omega, y = IGC1_Full_Homo.homogenizing_omega)) +
  geom_point() +
  geom_abline() +
  labs(title = "Homogenizing omega v.s. non-homogenizing omega", subtitle = "Homogenizing model") +
  xlab("non-homogenizing omega") +
  ylab("homogenizing omega")
```

Plot 2: In homogenizing model, comparing homogenizing omega and omega value.

\
\
\
\
\
\
\
\
\
\
\
\

## Tables

noIGC LogL: Log-likelihood of no IGC model (force $\tau=0$)

Xiang-noIGC llDiff: Log-likelihood difference between Xiang's original model and no IGC model

2omega-noIGC llDiff: Log-likelihood difference between 2 omega model (IGC_omega) and no IGC model

homoOmega(tau=0)-noIGC llDiff: Log-likelihood difference between no IGC homogenization model and no IGC model

homoOmega-noIGC llDiff: Log-likelood difference between homogenization model and no IGC model

```{r, echo=FALSE}
t = data.frame(IGC1_Full_forceTau$ll,
               IGC1_Full$ll - IGC1_Full_forceTau$ll,
               IGC1_Full_igcOmega$ll - IGC1_Full_forceTau$ll,
               IGC1_Full_forceTau_homoOmega$ll - IGC1_Full_forceTau$ll, 
               IGC1_Full_Homo$ll - IGC1_Full_forceTau$ll)
rownames(t) = rownames(IGC1_Full)
t %>%
  mutate_all(linebreak) %>%
  kable("latex", booktabs = T, escape = F,
        caption = 'Log-likelihood between different models',
        col.names = linebreak(c('noIGC\nlogL',
                                'Xiang\n-noIGC\nllDiff',
                                '2$\\omega$\n-noIGC\nllDiff',
                                'homoOmega\n(tau=0)\n-noIGC\nllDiff',
                                'homoOmega\n-noIGC\nllDiff'), align = "c")) %>%
  kable_styling(latex_options = c("striped"))
```

```{r, echo=FALSE}
kable(IGC1_Full_forceTau[,c(1,7,8,9)], "latex", booktabs = T, caption = "Xiang's original model with forcing tau to 0") %>%
  kable_styling(latex_options = c("striped"))

kable(IGC1_Full[,c(7,8,9)], "latex", booktabs = T, caption = "Xiang's original model") %>%
  kable_styling(latex_options = c("striped"))

kable(IGC1_Full_igcOmega[,c(7,8,9,10)], "latex", booktabs = T, caption = "2 omega model") %>%
  kable_styling(latex_options = c("striped"))

kable(IGC1_Full_forceTau_homoOmega[,c(1,7,8,9,10)], "latex", booktabs = T, caption = "Homogenization model with forcing tau to 0") %>%
  kable_styling(latex_options = c("striped"))

kable(IGC1_Full_Homo[,c(7,8,9,10)], "latex", booktabs = T, caption = "Homogenizing omega model") %>%
  kable_styling(latex_options = c("striped"))
```







