---
title: "Likelihood ratio test for new and old models"
author: "Yixuan Yang"
date: "9/24/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load library, include=FALSE}
rm(list = ls())
source("Functions.R")
library(tidyverse)
library(knitr)
```


```{r Read in data, include=FALSE}
# old data
IGC1_Full_noforce_noclock <- ListFunction("IGC1_Full_noforce_noclock", "Matrix") %>% t()
IGC1_Swap_noforce_clock <- ListFunction("IGC1_Swap_clock", "Matrix") %>% t()
IGC1_Unswap_noforce_clock <- ListFunction("IGC1_Unswap_clock", "Matrix") %>% t()
IGC2_Full_noforce_noclock <- ListFunction("IGC2_Full_noforce_noclock", "Matrix") %>% t()
IGC2_Swap_noforce_clock <- ListFunction("IGC2_Swap_clock", "Matrix") %>% t()
IGC2_Unswap_noforce_clock <- ListFunction("IGC2_Unswap_clock", "Matrix") %>% t()
# new data
IGC1_Full_spomega <- ListFunction("IGC1_Full_spomega", "Matrix") %>% t()
IGC1_Swap_clock_spOmega <- ListFunction("IGC1_Swap_clock_spOmega", "Matrix") %>% t()
IGC1_Unswap_clock_spOmega <- ListFunction("IGC1_Unswap_clock_spOmega", "Matrix") %>% t()
IGC2_Full_spomega <- ListFunction("IGC2_Full_spomega", "Matrix") %>% t()
IGC2_Swap_clock_spOmega <- ListFunction("IGC2_Swap_clock_spOmega", "Matrix") %>% t()
IGC2_Unswap_clock_spOmega <- ListFunction("IGC2_Unswap_clock_spOmega", "Matrix") %>% t()
```

# Hypothesis test
chi-square
```{r}
ll.test <- function(data.new, data.old, alpha=0.05){
  ll.new = data.new[,"ll"]
  ll.old = data.old[,"ll"]
  tau.new = data.new[,"tau"]
  tau.old = data.old[,"tau"]
  omega.new.point = data.new[,"omega"]
  omega.new.igc = data.new[,"IGC_omega"]
  omega.old = data.old[,"omega"]
  data1 = data.frame(ll.new, tau.new, omega.new.point, omega.new.igc)
  data2 = data.frame(ll.old, tau.old, omega.old)
  # merge two data sets to make them into same row number
  data = merge(data1, data2, by = 0) # by rownames
  data = data %>% drop_na() %>% data.frame(row.names = 1) %>% as.matrix() %>% round(4)
  # test part
  column1 = 2*(data[,"ll.new"]-data[,"ll.old"])
  column2 = 1-pchisq(column1, df=1) %>% round(4)
  column3 = column2 > alpha
  column3[column3 == TRUE] <- 'No'
  column3[column3 == FALSE] <- 'Yes'
  # combine 
  output = data.frame(data[,"tau.new"], data[,"tau.old"], data[,"omega.new.point"],
                      data[,"omega.new.igc"], data[,"omega.old"], column1, column2, column3)
  output = output[order(output[,6], decreasing = F),]
  colnames(output) = c("tau.new", "tau.old", "omega.new", "omega.igc", "omega.old", 
                       "2*(ll.new-ll.old)", "p-value", str_c("reject at ", alpha))
  return(output)
}
```

```{r, include=FALSE}
IGC1_Full_ll = ll.test(IGC1_Full_spomega, IGC1_Full_noforce_noclock)
IGC1_Swap_ll = ll.test(IGC1_Swap_clock_spOmega, IGC1_Swap_noforce_clock)
IGC1_Unswap_ll = ll.test(IGC1_Unswap_clock_spOmega, IGC1_Unswap_noforce_clock)
IGC2_Full_ll = ll.test(IGC2_Full_spomega, IGC2_Full_noforce_noclock)
IGC2_Swap_ll = ll.test(IGC2_Swap_clock_spOmega, IGC2_Swap_noforce_clock)
IGC2_Unswap_ll = ll.test(IGC2_Unswap_clock_spOmega, IGC2_Unswap_noforce_clock)
```

# Summary of the test
According to former investigation, we found Pillar 561,2321,4287,852,3278 have large branch length, and Pillar 3994 has large tau (small branch length). These data will need to be removed at the end, but we still keep them this time.

Some of the data from non-full gene data set are missing. Here we exclude them in the likelihood ratio test.

The test is to check if two times of the difference between the maximum likelihood of the two model follows a chi-square distribution.

**model1: with point mutation and IGC omega**

**model2: one omega**

# Table 1: Likelihood ratio test for 45 full gene data set
```{r}
kable(IGC1_Full_ll)
print(str_c("rejection number: ", sum(IGC1_Full_ll[,8]=='Yes')," of ", dim(IGC1_Full_ll)[1]))
```

# Table 2: Likelihood ratio test for reduced gene data set
```{r}
kable(IGC2_Full_ll)
print(str_c("rejection number: ", sum(IGC2_Full_ll[,8]=='Yes'), " of ", dim(IGC2_Full_ll)[1]))
```

# Swap test part
**species 1: green spotted puffer**

**species 2: zebra fish**

## Table 3: Likelihood ratio test for unswapped 45 full gene data set

```{r}
kable(IGC1_Unswap_ll)
print(str_c("rejection number: ", sum(IGC1_Unswap_ll[,8]=='Yes'), " of ", dim(IGC1_Unswap_ll)[1]))
```

## Table 4: Likelihood ratio test for unswapped reduced gene data set
```{r}
kable(IGC2_Unswap_ll)
print(str_c("rejection number: ", sum(IGC2_Unswap_ll[,8]=='Yes'), " of ", dim(IGC2_Unswap_ll)[1]))
```

## Table 5: Likelihood ratio test for swapped gene data set
```{r}
kable(IGC1_Swap_ll)
print(str_c("rejection number: ", sum(IGC1_Swap_ll[,8]=='Yes'), " of ", dim(IGC1_Swap_ll)[1]))
```

## Table 6: Likelihood ratio test for swapped reduced gene data set
```{r}
kable(IGC2_Swap_ll)
print(str_c("rejection number: ", sum(IGC2_Swap_ll[,8]=='Yes'), " of ", dim(IGC2_Swap_ll)[1]))
```


```{r, include=FALSE}
clean = function(data, n){
  name = str_c('MG', n)
  data = data[!(row.names(data) %in% name),]
  return(data)
}

n = c(561,2321,4287, 852, 3278, 3994)

```


```{r}
library(operators)
# missing data in reduced data set with old model
for( i in rownames(IGC2_Full_spomega)){
  if(i %!in% rownames(IGC2_Full_noforce_noclock)){
    print(i)
  }
}
# missing data in reduced data set with new model (swap test, unswapped)
for( i in rownames(IGC2_Unswap_noforce_clock)){
  if(i %!in% rownames(IGC2_Unswap_clock_spOmega)){
    print(i)
  }
}

```